---
title: "Conditional logit model"
bibliography: references.bib
date: "Last compiled on `r format(Sys.time(), '%B, %Y')`"
output: 
  html_document:
    css: tweaks.css
    toc:  true
    toc_float: true
    number_sections: true
    toc_depth: 1
    code_folding: show
    code_download: yes
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test3"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }
```


```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```


---  


To copy the code, click the button in the upper right corner of the code-chunks.

# clean up

```{r, results='hide'}
rm(list=ls())
gc()
```

<br>

# general custom functions

- `fpackage.check`: Check if packages are installed (and install if not) in R ([source](https://vbaliga.github.io/verify-that-r-packages-are-installed-and-loaded/))
- `fsave`: Function to save data with time stamp in correct directory
- `fload`: Function to load R-objects under new names

```{r, results='hide', message=FALSE, warning=FALSE}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file, location = "./data/processed/", ...) {
    if (!dir.exists(location))
        dir.create(location)
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, sep = "")
    print(paste("SAVED: ", totalname, sep = ""))
    save(x, file = totalname)
}


fload  <- function(fileName){
  load(fileName)
  get(ls()[ls() != "fileName"])
}
```

<br>

# necessary packages

- `tidyverse`
- `mlogit`
- `knitr`
- `kableExtra`

```{r, results='hide', message=FALSE, warning=FALSE}
packages = c("tidyverse", "mlogit", "knitr", "kableExtra")
fpackage.check(packages)
```

<br>

# load in mlogit data
```{r, eval=F}
#get todays date:
today <- gsub("-", "", Sys.Date())


abs_ml <- fload(paste0("./data/processed/",today,"abs_ml.RData"))
io_ml <- fload(paste0("./data/processed/",today,"io_ml.RData"))
```


<br>

# run model {.tabset .tabset-fade}

## ABS

```{r, eval=F}
m1 <- mlogit(choice ~ prestatie + kennis + gezellig + esteem | 0, abs_ml)
save(m1,file="./results/abs_m1.RData")

```

```{r}
ans_abs <- fload("./results/abs_m1.RData")
test <- summary(ans_abs)
res <- as_tibble(round(test$CoefTable, 3), rownames = "vars")

knitr::kable(res, digits=3, "html",  caption="Conditional logit model: preferences for sports partner attributes in ABS-survey") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

## I&O

Estimate models, same output in list and save the list

```{r, eval=F}
m1 <- mlogit(choice ~ prestatie + kennis + gezellig + esteem |0,io_ml)
#summary(m1)

save(m1,file="./results/io_m1.RData")
```


# {-}

# probe interaction {.tabset .tabset-fade}

Quick 'n dirty: subset data by gender and perform separate analyses. With I&O data.

```{r, eval=F}
io_ml


m1 <- mlogit(choice ~ prestatie + kennis + gezellig + esteem |0,io_ml)
#summary(m1)

save(m1,file="./results/io_m1.RData")
```

```{r}
ans_io <- fload("./results/io_m1.RData")
test <- summary(ans_io)
res <- as_tibble(round(test$CoefTable, 3), rownames = "vars")

knitr::kable(res, digits=3, "html",  caption="Conditional logit model: preferences for sports partner attributes in I&O-survey") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

