---
title: "ABS data"
bibliography: references.bib
date: "Last compiled on `r format(Sys.time(), '%B, %Y')`"
output: 
  html_document:
    css: tweaks.css
    toc:  true
    toc_float: true
    number_sections: true
    toc_depth: 1
    code_folding: show
    code_download: yes
---

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test3"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }
```


```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```


---  


To copy the code, click the button in the upper right corner of the code-chunks.

# clean up

```{r, results='hide'}
rm(list=ls())
gc()
```

<br>

# general custom functions

- `fpackage.check`: Check if packages are installed (and install if not) in R ([source](https://vbaliga.github.io/verify-that-r-packages-are-installed-and-loaded/))
- `fsave`: Function to save data with time stamp in correct directory
- `fload`: Function to load R-objects under new names

```{r, eval=FALSE}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file, location = "./data/processed/", ...) {
    if (!dir.exists(location))
        dir.create(location)
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, sep = "")
    print(paste("SAVED: ", totalname, sep = ""))
    save(x, file = totalname)
}


fload  <- function(fileName){
  load(fileName)
  get(ls()[ls() != "fileName"])
}
```

<br>

# necessary packages

- `tidyverse`
- `mlogit`

```{r, eval=FALSE}
packages = c("tidyverse", "mlogit")
fpackage.check(packages)
```

<br>

# load in raw data

I load in raw, non-public data of the second wave of the survey project *Sports & Networks* [Franken et al., -@data2022]. This wave contains the DCE data. The data were deposited in [DANS EASY](link), but are not publicly accessible. 

```{r,eval=F}
df <- read.csv("./rawdata/results-survey868275.csv")
```

<br>

# data wrangling

```{r,eval=F}
# first, delete fake response;
fakeid <- which(df$email  %in% "rob.franken@ru.nl")
df <- df[-fakeid,]

# add id 
df$id <- 1:nrow(df)

# to long format

# each list element contains the choices for each set. thus presatielist[[2]][3] is the
# 'presatie'-attribute of person 3, in choice set 

prestatielist <- list()
prestatielist[[1]] <- c("attribute_16", "attribute_17", "attribute_18")
prestatielist[[2]] <- c("attribute_19", "attribute_20", "attribute_31")
prestatielist[[3]] <- c("attribute_32", "attribute_33", "attribute_34")

kennislist <- list()
kennislist[[1]] <- c("attribute_35", "attribute_36", "attribute_37")
kennislist[[2]] <- c("attribute_38", "attribute_39", "attribute_40")
kennislist[[3]] <- c("attribute_41", "attribute_42", "attribute_43")

gezelliglist <- list()
gezelliglist[[1]] <- c("attribute_44", "attribute_45", "attribute_46")
gezelliglist[[2]] <- c("attribute_47", "attribute_48", "attribute_49")
gezelliglist[[3]] <- c("attribute_50", "attribute_51", "attribute_52")

esteemlist <- list()
esteemlist[[1]] <- c("attribute_53", "attribute_54", "attribute_55")
esteemlist[[2]] <- c("attribute_56", "attribute_57", "attribute_58")
esteemlist[[3]] <- c("attribute_59", "attribute_60", "attribute_61")

chosen <- c("CS1", "CS2", "CS3")


for (set in 1:3) {

    # to long format
    
  for (choice in 1:3) {

  
        data <- as.data.frame(df)
        data$set <- set
        data$options <- choice
        data$prestatie <- data[, names(data) == prestatielist[[set]][choice]]
        data$kennis <- data[, names(data) == kennislist[[set]][choice]]
        data$gezellig <- data[, names(data) == gezelliglist[[set]][choice]]
        data$esteem <- data[, names(data) == esteemlist[[set]][choice]]
        data$chosen <- data[, names(data) == chosen[set]]
        data <- data[, names(data) %in% c("id", "set", "options", "prestatie", "kennis", "gezellig",
            "esteem", "chosen")]
        if (set == 1 & choice == 1) {
            df_long <- data
        } else {
            df_long <- rbind(df_long, data)
        }
    }
}

# order
df_long <- df_long[order(df_long$id, df_long$set, df_long$options), ]

# define the choice
df_long$choice <- (df_long$chosen == c("Persoon 1", "Persoon 2", "Persoon 3")[df_long$options])
```

<!-- @RF: to-do, match ego-level covariates for descriptive statistics and moderational analyses (gender, sports history, age, ...) --->

<br>

# make ready for mlogit

```{r, eval=F}
# first we need an id_alt (identification of alternatives), thus specific combi of facets
{
    df_long$id_alt <- as.numeric(as.factor(trimws(df_long$prestatie))) * 1000 + as.numeric(as.factor(trimws(df_long$kennis))) *
        100 + as.numeric(as.factor(trimws(df_long$gezellig))) * 10 + as.numeric(as.factor(trimws(df_long$esteem)))

}

# we also need an id for each specific choice set, thus combi of respnr and set
df_long$id_choice <- df_long$id * 10 + df_long$set

length(unique(df_long$id))
length(unique(df_long$id_choice))
sum(df_long$choice)#no missings on choice. !be aware these need to be deleted for mlogit. 

# need to delete id_choice where no choice is made

# do we have duplicates?
df_long %>%
    select(c("id_choice", "id_alt")) -> df_sel
dups <- duplicated(df_sel)
sum(dups)

# and with no duplicates and no choice sets without choices, this should work 
df_ml <- mlogit.data(df_long, choice = "choice", id.var = "id", chid.var = "id_choice", alt.var = "id_alt",
    shape = "long")

```

<br>

# save data object
```{r, eval=F}
fsave(df_ml, "abs_ml.RData")
```
<br>

---

## References


